The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-10-16 paper/Analisis.tex
changed from MNRAS to ApJ

\documentclass[usenatbib]{mn2e} 
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{epsfig} 
\usepackage{float} 
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}
\bibliographystyle{mn2e}
% To highlight comments 
\usepackage{color}
\definecolor{red}{rgb}{1,0.0,0.0}
\newcommand{\red}{\color{red}}
\definecolor{blue}{rgb}{0.1,0.3,0.9}
\newcommand{\blue}{\color{blue}}

\usepackage[normalem]{ulem}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

\title[Rotation in the Lyman-$\alpha$ line]{The impact of gas bulk
  rotation on the Lyman-$\alpha$ line} 
\author[Garavito-Camargo, Forero-Romero \& Dijkstra]{
\parbox[t]{\textwidth}{\raggedright 
  Nicolas Garavito-Camargo$^{1}$,
  Jaime E. Forero-Romero$^{1}$ and 
  Mark Dijkstra$^2$
}
\vspace*{6pt}\\
$^{1}$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^2$
}
\maketitle

\begin{abstract}
Rotation is present in the gas kinematics of galaxies up to the
highest redshifts. In this paper we present for the first time 
radiative transfer calculations that show the impact of gas bulk rotation on
the morphology of the Lyman $\alpha$ line. To this end we model a
galaxy as an homogeneous sphere composed as an homogeneous mixture of
dust and hydrogen at a constant temperature. These spheres have a
solid-body rotation with linear velocities at the surface in the range
$0-300$ \kms and neutral hydrogen optical depths in the range
$\tau_{\rm H}=10^{5}-10^{7}$. We consider radiation sources both in
the center of the rotating cloud and also homogeneously distributed in
the volume. We find that higher rotational velocities increase the
width of each peak in the outgoing line profile while it also
increases the amount of Lyman alpha photons escaping in the line
center. This trends makes that for high rotational velocities and
large Hydrogen optical depths the double peak of the line tends to be
erased an be replaced by a single peak a the line center. This trend
is more pronounced for radiation sources homogeneously
distributed. Concerning the escape fraction we find that rotation does
not have any effect, provided that all the sources are centrally
emitted. However, in the case of homogeneously emittedsources we
measure an increase of about a factor of $2$ in the escape  fraction
for higher rotational velocity values. This work shows that gas bulk
rotation has a non negligible impact on the shape of the Lyman
$\alpha$ line.   
\end{abstract}
\begin{keywords}
galaxies: high-redshift - galaxies: star formation - line: formation
\end{keywords}

@@

\begin{subequations}
\begin{align}
    v_{x}=-\dfrac{y}{R}V_{\rm max}, \label{subeq1}\\
    v_{y}=\dfrac{x}{R}V_{\rm max}, \label{subeq2}\\
    v_{z}=0, \label{subeq3}
\end{align}
\end{subequations}
%
where $R$ is the radius of the sphere and $V_{\rm max}$ is the linear
velocity at the sphere's surface. The minus/plus sign in the
$x$/$y$-component of the velocity indicates the direction of
rotation. In this case we take the angular velocity in the same
direction as the $\hat{k}$ unit vector. With these definitions we can
write the angular velocity as $\omega=V_{\rm max}/R$.

@@

\bibliography{references}

>>>>

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJ}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}
%\bibliographystyle{apj}
\usepackage{graphics}
\usepackage[dvips]{epsfig}

%\documentclass[usenatbib]{mn2e} 
%\usepackage{amsmath} 
%\usepackage{amssymb} 
%\usepackage{graphics}
%\usepackage{graphicx}
%\usepackage{epsfig} 
%\usepackage{float} 
%\def\be{\begin{equation}}
%\def\ee{\end{equation}}
%\def\ba{\begin{eqnarray}}
%\def\ea{\end{eqnarray}}
%\bibliographystyle{mn2e}
% To highlight comments 
%\usepackage{color}
%\definecolor{red}{rgb}{1,0.0,0.0}
%\newcommand{\red}{\color{red}}
%\definecolor{blue}{rgb}{0.1,0.3,0.9}
%\newcommand{\blue}{\color{blue}}

%\usepackage[normalem]{ulem}
%\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
%\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
%\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
%\newcommand{\LCDM}{$\Lambda$CDM~}
%\newcommand{\beq}{\begin{eqnarray}}  
%\newcommand{\eeq}{\end{eqnarray}}

%\newcommand{\zz}{$z\sim 3$} 
%\newcommand{\apj}{ApJ}  
%\newcommand{\apjs}{ApJS}  
%\newcommand{\apjl}{ApJL}  
%\newcommand{\aj}{AJ}  
%\newcommand{\mnras}{MNRAS}  
%\newcommand{\mnrassub}{MNRAS accepted}  
%\newcommand{\aap}{A\&A}  
%\newcommand{\aaps}{A\&AS}  
%\newcommand{\araa}{ARA\&A}  
%\newcommand{\nat}{Nature}  
%\newcommand{\physrep}{PhR}
%\newcommand{\pasp}{PASP}    
%\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

@@

\title{The impact of gas bulk rotation on the Lyman-$\alpha$ line} 
\author{
  Nicolas Garavito-Camargo\altaffilmark{1},
  Jaime E. Forero-Romero\altaffilmark{1},
  Mark Dijkstra\altaffilmark{2}
}

\altaffiltext{1}{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\altaffiltext{2}{--}

\begin{abstract}
Rotation is present in the gas kinematics of galaxies up to the
highest redshifts. In this paper we present for the first time 
radiative transfer calculations that show the impact of gas bulk rotation on
the morphology of the Lyman $\alpha$ line. To this end we model a
galaxy as an homogeneous sphere composed as an homogeneous mixture of
dust and hydrogen at a constant temperature. These spheres have a
solid-body rotation with linear velocities at the surface in the range
$0-300$ \kms and neutral hydrogen optical depths in the range
$\tau_{\rm H}=10^{5}-10^{7}$. We consider radiation sources both in
the center of the rotating cloud and also homogeneously distributed in
the volume. We find that higher rotational velocities increase the
width of each peak in the outgoing line profile while it also
increases the amount of Lyman alpha photons escaping in the line
center. This trends makes that for high rotational velocities and
large Hydrogen optical depths the double peak of the line tends to be
erased an be replaced by a single peak a the line center. This trend
is more pronounced for radiation sources homogeneously
distributed. Concerning the escape fraction we find that rotation does
not have any effect, provided that all the sources are centrally
emitted. However, in the case of homogeneously emittedsources we
measure an increase of about a factor of $2$ in the escape  fraction
for higher rotational velocity values. This work shows that gas bulk
rotation has a non negligible impact on the shape of the Lyman
$\alpha$ line.   
\end{abstract}

\begin{keywords}
{galaxies: high-redshift,galaxies: star formation, line: formation}
\end{keywords}

@@

\begin{equation}
    v_{x}=-\frac{y}{R}V_{\rm max}, \label{subeq1}
\end{equation}
\begin{equation}
    v_{y}=\frac{x}{R}V_{\rm max}, \label{subeq2}
\end{equation}
\begin{equation}
    v_{z}=0, \label{subeq3}
\end{equation}
%
where $R$ is the radius of the sphere and $V_{\rm max}$ is the linear
velocity at the sphere's surface. The minus/plus sign in the
$x$/$y$-component of the velocity indicates the direction of
rotation. In this case we take the angular velocity in the same
direction as the $\hat{k}$ unit vector. With these definitions we can
write the angular velocity as $\omega=V_{\rm max}/R$.

@@

\bibliographystyle{apj}
\bibliography{references}
